---
title: "HW 9 - Multiple Regression Model - OLS Asymptotics"
author: "Marcelo Pignatari"
date: "11/02/2021"
output:
  word_document: default
  pdf_document: default
---

\tableofcontents


\newpage

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen=999)
```


```{r include=FALSE}
library(wooldridge)
library(stargazer)
library(dplyr)
library(car)
library(ggplot2)
```

# Problems

## Question 1

As risktol should be positively correlated with pctstck (higher tolerance for risk, higher investment in stock market), so \beta_2>0 and $\tilde{\beta_1}$ has a positive inconsistency.

Since $plim\tilde{\beta_1}=\beta_1+\beta_2\delta_1$, $plim\tilde{\beta_1}=plim\beta_1+plim\beta_2\delta_1=\beta_1+plim\beta_2\delta_1$
As $\beta_2>0$ and $\delta_1>0$ → $plim\beta_2\delta_1>0$.

## Question 2

### (i)

$$ H_0: \beta_3=0$$
$$ H_1: \beta_3\neq0$$

### (ii)

$\beta_1>0\,and\,\beta_2>0$. It expected that most populated and richer cities has highest rent rates.

### (iii)

The correct statement is: "A 10% increase in population is associated with about 0.66% increase in rent".

### (iv)

As t-stat = 56/17 = **`r 56/17`** > 3, and the tstat of $\alpha$=1%(critical value) is **`r qt(0.995,60)`**. Since t>c we reject $H_0$, i.e. we reject the assuption that that the student body relative to the population has no ceteris paribus effect on monthly rents.

## Question 3

### (i)

$$Var(\hat{\beta_1}-3\hat{\beta_2})=E[\hat{\beta_1}-3\hat{\beta_2}+E[\hat{\beta_1}-3\hat{\beta_2}]^2]=E[((\hat{\beta_1}-E(\hat{\beta_1})+(-3\hat{\beta_2}-E[3\hat{\beta_2}))^2]=E[(\hat{\beta_1}-E(\hat{\beta_1})^2+(-3\hat{\beta_2}-E[3\hat{\beta_2}))^2+2(\hat{\beta_1}-E(\hat{\beta_1})(-3\hat{\beta_2}-E[3\hat{\beta_2}))$$
$$=E[(\hat{\beta_1}-E(\hat{\beta_1})^2]+E[(-3\hat{\beta_2}-E[-3\hat{\beta_2}])^2]+2(\hat{\beta_1}-E(\hat{\beta_1})(-3\hat{\beta_2}-E[-3\hat{\beta_2}))=Var(\hat{\beta_1})+Var(-3\hat{\beta_2})+2Cov(\hat{\beta_1},-3\hat{\beta_2})=Var(\hat{\beta_1})+9Var(\hat{\beta_2})-6Cov(\hat{\beta_1},-3\hat{\beta_2})$$
As $Var(\hat{\beta_1}-3\hat{\beta_2}) = Var(\hat{\beta_1})+9Var(\hat{\beta_2})-6Cov(\hat{\beta_1},-3\hat{\beta_2})$, then $se(\hat{\beta_1}-3\hat{\beta_2})=\sqrt{Var(\hat{\beta_1})+9Var(\hat{\beta_2})-6Cov(\hat{\beta_1},-3\hat{\beta_2})}$


### (ii)

$$t=\frac{\hat{\beta_1}-3\hat{\beta_2}-1}{\sqrt{Var(\hat{\beta_1})+9Var(\hat{\beta_2})-6Cov(\hat{\beta_1},-3\hat{\beta_2})}}$$

### (iii)


Given $\theta_1=\beta_1-3\beta_2$, we have that $\beta_1=\theta_1+3\beta_2$. Since $y=\beta_0+\beta_1x_1+\beta_2x_2+\beta_3x_3+u$, we can substitute $\beta_1$ in terms of $\theta$ on the equation:
$y=\beta_0+(\theta_1+3\beta_2)x_1+\beta_2x_2+\beta_3x_3+u$
$y=\beta_0+\theta_1x_1+3\beta_2x_1+\beta_2x_2+\beta_3x_3+u$
$y=\beta_0+\theta_1x_1+\beta_2(3x_1+x_2)+\beta_3x_3+u$

$y=\hat{\beta_0}+\hat{\theta_1}x_1+\hat\beta_2(3x_1+x_2)+\hat\beta_3x_3$

With this transformation we can obtain $\hat{\theta_1},se(\hat{\theta_1})$ and test said hypothesis.

# Question 4 

## (i) 


```{r}
qt(0.9995,86) # 0,1% cv
qt(0.995,86) # 1% cv
qt(0.975,86) # 5% cv
qt(0.95,86) # 10% cv
round(-14.47/16.27,2)
round((0.976-1)/0.049,2)
```

The first hypothesis test is $H_0:\beta_0=0$ and $H_1:\beta_0\neq0 $. 
Since t-stat = |**`r round(-14.47/16.27,2)`**|<|1.66| t-stat of 10% critical value (the highest standard tolerance) we can't reject $H_0$.

The second hypothesis test is $H_0:\beta_1=1$ and $H_1:\beta_1\neq1$

Since t-stat = |**`r round((0.976-1)/0.049,2)`**|<|1.66| t-stat of 10% critical value (the highest standard tolerance) we can't reject $H_0$. Its even less significant.

## (ii) 

```{r}
((209448.99-165644.51)/2)/(165644.51/86)
qf(0.999,2,86)
```


The hypothesis test is $H_0:\beta_0=0\, and \,\beta_1=1$ $H_1: \beta_0\neq0\, and \,\beta_1\neq1$.

$$F=\frac{(SSR_r-SSR_{ur})/q}{SSR_{ur}/(n-k-1)} = \frac{(209,448.99-165,644.51)/2}{165,644.51/86} = 11.3713$$
Since |11.37|>7.5, we reject the null hypothesis overwhemingly even with critical F value at 0.1%.

## (iii)

```{r}

qf(0.9,3,83)

((0.829-0.820)/3)/((1-0.829)/83)
  
```
  
$$F=\frac{(R^2_{ur}-R^2_r)/q}{(1-R^2_{ur})/(n-k-1)} = \frac{(0.829-0.820)/3}{(1-0.829)/83} = 1.46$$
Since |1.46|<2.15, we can't reject the null hypothesis even with tolerance F value at 10%.

### (iv)

We have heteroskedasticity violating MLR5 assumption. This means that we cannot estabilish a reliable distribution of probabilities for the coefficients and therefore we cannot have a reliable hypothesis testing. 

## Computer Exercises

## Question 5 

### (i)

```{r}
data("k401ksubs")
#head(k401ksubs)

d2<-filter(k401ksubs,fsize==1)
count(d2)
head(d2)
```
There are 2017 single person households in the dataset.

### (ii) 

```{r}
mrm1<-lm(nettfa~inc+age,d2)
summary(mrm1)
stargazer(mrm1,type = "text")
```

$$\widehat{nettfa}=-43.040+0.799inc+0.843age$$
$R^2=0.119, n=2,017$

Any additional thousand (1000) dollars in annual income predicts - on average and holding everything else constant - an increase in 799 dollars in net total financial assets.
Any additional age an individual has predicts - on average and holding everything else constant - an increase in 843 dollars in net total financial assets. 

Regarding the sign of the slope coefficients I am not surprised, it is expected that more income generates more wealth and the older the person is, more time she has to earn money and save money. 
About the magnitude it is kind of surprising because it means that the overral population has a satisfactory capacity to save money in relation to their income and through the time.     

### (iii)

The intercept says that a person with no income that and just born (age 0) has a negative wealth of 43 (measure in 1000s) dollars. That is, people are born with a small debt. If you try hard maybe you can find a meaning, but is not interesting. 


### (iv) 

The hypothesis test is $H_0:\beta_2=1$ against $H_1: \beta_2<1$.
```{r}
t<-(0.843-1)/0.092
t
df<-2017-3
pt(t,df)
```

P-value is 0.044, which means that the smallest significance level that  $H_0$ is still rejected is 4.4%. Therefore, we can't reject the null hypothesis at 1% significance level. 

### (v)

```{r}
mrm2<-lm(nettfa~inc,d2)
stargazer(mrm2,type = "text")
mrm3<-lm(inc~age,d2)
stargazer(mrm3,type = "text")
```

No, is very close. That must be because inc is not highly correlated with age, hence when we drop age the coefficient of income doesn't suffer much bias. 

## Question 6

### (i)

```{r}
data("discrim")
#head(discrim)
mrm4<-lm(log(psoda)~prpblck+log(income)+prppov,discrim)
stargazer(mrm4,type = "text")

```

$$\widehat{log(psoda)}=0.073prpblck+0.137log(income)+0.380prppov$$
$$R^2=0.087, n=401$$

$\beta_1$ is statistically different from zero at the 5% level but not at the 1% of tolerance. 

### (ii)

```{r}
cor(discrim$lincome,discrim$prppov,use="complete.obs")
summary(mrm4)
```

The correlation between log(income) and prppov is -0.838. log(income) is statiscally significant at almost 0% tolerance level,  prppov is statiscally significant at 0.1% significance level. The p-values are 0.00000048 and 0.0044 respectively.

### (iii)


```{r}
H0<-c("prpblck=0","prppov=0")
linearHypothesis(mrm4,H0)

```

For sure, its highly significant almost in zero level. P-value is almost zero. 


### (iv)

```{r}
mrm5<-lm(log(psoda)~prpblck+log(income)+prppov+log(hseval),discrim)
stargazer(mrm5,type = "text")
summary(mrm5)
```

log(hseval) coefficient means that an additional 1% unit in median housing value at a zipcode predicts - on average and holding everything else constant - an increase in 
The p-value of log(hseval) is overwhelmingly different from zero, its value is practically zero for rejecting H0.

### (v)

In order to test the hypothesis $H_0:\beta_1=\beta_3$, it's necessary set $\theta_1=\beta_1-\beta_3$ and rewrite the model as:

$$
log(soda)=\beta_0+(\theta_1+\beta_3)prpblck+\beta_2lincome+\beta_3prppov+\beta_4lhseval+u
$$
$$
log(soda)=\beta_0+\theta_1prpblck+\beta_2lincome+\beta_3(prppov+prpblck)+\beta_4lhseval+u
$$
```{r}
mrm6<-lm(lpsoda ~ prpblck+lincome+I(prppov+prpblck)+lhseval,discrim)
summary(mrm6)
```




Since the coefficient $\theta=0.04$, and $\theta=\beta_1-\beta_3$, we need to test the hypothesis $H_0:\theta=0$
```{r}
tstat6<-(mrm6$coefficients[2]-0)/0.15279
ptst<-(1-pt(tstat6,396))*2
ptst
```

Since the p-value for the two tail test, is 0.766, we fail to reject the null hypothesis that $H_0:\theta=0$ with $\alpha=1%$. In other words, we fail to reject that the effect of prpblck on psoda is the same as the effect of prppov on psoda.




