---
title: "HW 8 - Multiple Regression Model - OLS Asymptotics"
author: "Marcelo Pignatari"
date: "11/02/2021"
output:
  word_document: default
  pdf_document: default
---

\tableofcontents


\newpage

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen=999)
```


```{r include=FALSE}
library(wooldridge)
library(stargazer)
library(dplyr)
library(car)
library(ggplot2)
```

# Problems

## Question 1

As risktol should be positively correlated with pctstck (higher tolerance for risk, higher investment in stock market), so $\beta_2>0$ and $\tilde{\beta_1}$ has a positive inconsistency.

Since $plim\tilde{\beta_1}=\beta_1+\beta_2\delta_1$, 
As $\beta_2>0$ and $\delta_1>0$ → $plim\tilde{\beta_1}= \beta_1 + \beta_2\delta_1>\beta_1$.

## Question 2

### (i)

```{r}
data(wage1)
mrm1<-lm(wage~educ+exper+tenure,wage1)
#stargazer(mrm1,type = "text")
res1<-residuals(mrm1)
hist(res1)

```


### (ii)

```{r}
mrm2<-lm(log(wage)~educ+exper+tenure,wage1)
stargazer(mrm2,type = "text")
res2<-residuals(mrm2)
hist(res2)
lines(density(res2))
```

### (iii)

The residuals of the log-level mode seem to have a distribution closer to normal. So MLR 6 is closer to being satisfied on this model.  Level-level is more right skewed and has more residuals far from the mean. 

## Question 3

### (i)

```{r}
data(econmath)
head(econmath)
range(econmath$score)
e<-filter(econmath, !is.na(econmath$actmth))
range(e$scor)
range(econmath,na.rm = T)
```

The smallest value is 19.53 the largest value is 98.44.  The smallest value in the sample is 20.31, the largest value is 98.44.

*I'm a little confused by this question, if the second question is what is the range of values in the entire database the answer is 0 to 98.44.

### (ii)

```{r}
mrm2<-lm(score~colgpa+actmth+acteng,econmath)
stargazer(mrm1,mrm2,type = "text")
varb<-summary(mrm1)$coefficients[,2]

```

MLR6 is not holding for the error term because the variable score is limited to the interval [0,100] and a normally distributed random variable cannot have limits. If the dependent variable is not normallt distributed, a linear combination of the dependent variable (i.e. the error terms) can not be normally distributed either.

However, with a large sample size we can rely on the assymptotics properties of the estimate bt the central limit theorem. Then, in this case, for n=814 we can evaluate $H_0:\beta_3=0$.

```{r}
hist(resid(mrm2),freq=F)


```



### (iii) 



```{r}
summary(mrm2)


```

The p-value associated with $H_0:\beta_3=0$ is 0.641, the t-stat is 0.466 . We can defend this findings arguing that besides the fact that the error seems not have a normal distribution, the sample has a large number of obervations. Therefore we can assume that the coefficients have consistency and by this we can rely on the p-value.
