---
title: "Problem Set 1"
author: "Marcelo Pignatari"
date: "01/26/2022"
output:
  word_document: default
  pdf_document: default
---

\tableofcontents


\newpage

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen=999)
```


```{r include=FALSE}
library(astsa)
library(tibble)
library(knitr)
library(xts)
library(xtable)
```

# R programming exercises

## 1.
```{r}
sqrt(729)

```

## 2.
```{r}
a<-1947.0

```

## 3.
```{r}
b<-c(1,2,3,4,5,6)
class(b)

```

## 4.
```{r}
c<-c(1,"a",2,"b")

```
### a)
```{r}
class(c)

```
### b)
```{r}
length(c)

```
### c)
```{r}
c[c(2,3)]

```

## 5.
```{r}
d<-c(1,2,NA,4,5,6,NA,NA,NA,10)
d[-c(3,7,8,9)]
d[!is.na(d)]

```

## 6.
```{r}
e<-function(x){exp(x)*cos(x)}
f<- c(e(seq(3,6,0.1)))
f
x<-seq(3,6,by=.1)
f1<-exp(x)*cos(x)
f1
```

## 7.
```{r}
g<-function(x){x^3+4*x^2} 
h<-sum(g(seq(10,100,1)))
g1<-seq(10,100,1)
h1<-sum(g1^3+4*g1^2)
h
h1
```
## 8.
```{r}
x<- sample(0:999,size=250,replace = TRUE)
y<- sample(0:999,size=250,replace = TRUE)

```

### a)
```{r}
y[y>600]

```
### b)
```{r}
length((y[y>=max(y)-200]))

```

### c)
```{r}
e<-sqrt(abs(x-mean(x)))
e

```

### d)
```{r}
x1<-c(0,x) 
y1<-c(y,0)
g4<-(y1-x1) 
g4

i<-y[y=seq(2,length(y),1)]-x[x=seq(1,length(x)-1,1)]
i

```

## 9.
```{r}
coeffs<-sample(-20:20,size=3,replace=TRUE)

```

### a)
```{r}
print(coeffs)

length(coeffs)

```

### b)
```{r}
q<-seq(-3,3,length.out=200)
q

```

### c)
```{r}
z<-function(x){coeffs[1]+coeffs[2]*x+coeffs[3]*x^2}
z(q)
z1<-coeffs[1]+coeffs[2]*q+coeffs[3]*q^2

```

### d)
```{r}
plot(q,z(q))

```

## 10.
$$1/8-1+6-1=33/8$$

## 11.
```{r}
s<-matrix(rep(c(10,-5,10),15),nrow=15,ncol=3,byrow=TRUE)
s

s[,3]<-s[,1]+s[,2]
s

```

## 12.
```{r}
t<-function(x1,x2){x1+x2}
t
t(7,8)

```

## 13.
```{r}
x1<-c(1,2,2,3,4,5,6,2,2,2,2)
beta<-function(x1,x2){sum(x1==x2)}
beta(x1,2)

length(which(x1==2))
```

## 14

```{r}
v<-c(1,2,3,6)
csqt<-function(v){return(v^seq(1:length(v)))}
csqt(v)

```
## 15
```{r}
ffunc<-function(v){return(ifelse(v<0,v^2+2*v+3,ifelse(v<2,v+3,v^2+4*v-7)))}
inp<-seq(-3,3,length=100)

asd<-ffunc(inp)

plot(inp,asd)

```

# Theory

## Problem 1

Given $E(Y_t) = 0$ and $Var(Y_t)=\sigma^2$ for any t in T. 
Because $Var(Y_t)$ is the second moment of the series we have: $Var(Y_t)=E[(Y_t-E(Y_t))^2]=E[(Y_t-0)^2]=E[Y_t]^2$

## Problem 2

$$\gamma_Y(s,t)=E[(Y_s-E(Y_s))(Y_t-E(Y_t))]=E[(Y_s-\mu_s)(Y_t-\mu_t)]=E[(Y_sY_t-Y_s\mu_t-Y_t\mu_s+\mu_s\mu_t)]=E[Y_sY_t]-E[Y_s]\mu_t-E[Y_t]\mu_s+E[\mu_s\mu_t)]=E[Y_sY_t]-\mu_s\mu_t-\mu_s\mu_t+\mu_s\mu_t=E(Y_sY_t) - \mu_s\mu_t$$

## Problem 3

```{r}

tblnm<-c("$y_t$","$y_{t−1}$","$y_{t−2}$","$y_{t−3}$","$\\hat\\mu_Y(t)$","$y_t−\\hat\\mu_Y(t)$","$(y_t−\\hat\\mu_Y(t))^2$","$y_{t-1}-\\hat\\mu_Y(t)$","$(y_t-\\hat\\mu_Y(t))(y_{t-1}-\\hat\\mu_Y(t))$")

x<-c(112,118,132,129,121,135,148,148,136,119,104,118)
#xtbl<-cbind(rep(1949,12),c(112,118,132,129,121,135,148,148,136,119,104,118))
xrnm<-c("Jan 1949","Feb 1949","Mar 1949","Apr 1949","May 1949","Jun 1949","Jul 1949","Aug 1949","Sep 1949","Oct 1949","NOv 1949","Dec 1949","Sum")

xtbl<-ts(x,start=c(1949,1),frequency = 12)
mu<-mean(xtbl)

xtbl1<-cbind(xtbl,lag.xts(xtbl),lag.xts(xtbl,k=2),lag.xts(xtbl,k=3),rep(mu,12),xtbl-mu,(xtbl-mu)^2,lag.xts(xtbl)-mu,(xtbl-mu)*(lag.xts(xtbl)-mu))

xtbl2<-as.data.frame(rbind(xtbl1,colSums(xtbl1,na.rm = T)))
rownames(xtbl2)<-xrnm
colnames(xtbl2)<-tblnm

kable(xtbl2, booktab=T,digits=2)

cov1<-sum(xtbl2[-NROW(xtbl2),9],na.rm = T)

v<-sum(xtbl2[2:(NROW(xtbl2)-1),7],na.rm=T)

v1<-sum(xtbl2[2:(NROW(xtbl2)-1),8]^2,na.rm = T)


```

Sample variance: $\frac{\sum ( y_t − \hat{\mu}_Y(t))^2}{T-1} = \frac{`r round((xtbl2[NROW(xtbl2),7]),2)`}{`r (sum(!is.na(xtbl2[,7]))-2)`}$= **`r (round(xtbl2[NROW(xtbl2),7]/(sum(!is.na(xtbl2[,7]))-2),2))`**  

Sample autocovariance: $\frac{\sum_{t=2}^T (y_t-\hat\mu_Y(t))(y_{t-1}-\hat\mu_Y(t))}{T-1} = \frac{`r round((xtbl2[NROW(xtbl2),9]),2)`}{`r (sum(!is.na(xtbl2[,9]))-2)`}$=
**`r (round(xtbl2[NROW(xtbl2),9]/(sum(!is.na(xtbl2[,9]))-2),2))`**   

Sample autocorrelation: $\frac{\frac{\sum_{t=2}^T (y_t-\hat\mu_Y(t))(y_{t-1}-\hat\mu_Y(t))}{T-1}}{\sqrt{(\frac{\sum_{t=2}^T (y_t−\hat\mu_Y(t))^2 }{T-1})(\frac{\sum_{t=2}^T (y_{t-1}−\hat\mu_Y(t))^2}{T-1})}}$ = $\frac{`r round(cov1,1)`}{`r round(sqrt(v*v1),1)`}$ = 
**`r round(cov1/sqrt(v*v1),3)`**.  




